---
title: "SIA sample selection"
author: "Jonah Bacon"
date: '2023-01-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(knitr)
library(kableExtra)

lw.dat <- read.csv("data/Sample_LW.csv")
lw.dat <- lw.dat %>% unite(spp_ID, species:ID, sep = "_", remove = FALSE)
lw.dat$species <- as.factor(lw.dat$species)
lw.dat$date_collected <- as.Date(lw.dat$date_collected, format = "%d-%b-%Y")
lw.dat$date_processed <- as.Date(lw.dat$date_processed, format = "%d-%b-%Y")
lw.dat$net_ID <- as.factor(lw.dat$net_ID)
lw.dat$weight_g <- as.numeric(lw.dat$weight_g)
lw.dat$gear_type <- as.factor(lw.dat$gear_type)
lw.dat$lg_group <- floor(lw.dat$length_mm/25)
str(lw.dat)

lw21 <- lw.dat %>% filter(date_collected <= "2021-12-31")
lw22 <- lw.dat %>% filter(date_collected >= "2022-01-01")
```
```{r miscellaneous_unused, include=FALSE}
# hist(arcs22$length_mm, breaks = c(seq(100,400,25)))

# length.dist.df <- lw22 %>% 
#   group_by(species) %>% 
#   summarise(
#     "<100" = sum(length_mm < 100),
#     "100-124" = sum(length_mm >= 100 & length_mm < 125),
#     "125-149" = sum(length_mm >= 125 & length_mm < 150),
#     "150-174" = sum(length_mm >= 150 & length_mm < 175),
#     "175-199" = sum(length_mm >= 175 & length_mm < 200),
#     "200-224" = sum(length_mm >= 200 & length_mm < 225),
#     "225-249" = sum(length_mm >= 225 & length_mm < 250),
#     "250-274" = sum(length_mm >= 250 & length_mm < 275),
#     "275-299" = sum(length_mm >= 275 & length_mm < 300),
#     "300-324" = sum(length_mm >= 300 & length_mm < 325),
#     "325-349" = sum(length_mm >= 325 & length_mm < 350),
#     "350-374" = sum(length_mm >= 350 & length_mm < 375),
#     "375-399" = sum(length_mm >= 375 & length_mm < 400),
#     "400-424" = sum(length_mm >= 400 & length_mm < 425),
#     "425-449" = sum(length_mm >= 425 & length_mm < 450),
#     "450-474" = sum(length_mm >= 450 & length_mm < 475),
#     "475-499" = sum(length_mm >= 475 & length_mm < 500),
#     ">500" = sum(length_mm >= 500)
#     )
# 
# set.seed(101)
# sample(arcs22$ID, size = 20, replace = FALSE)
```

# Randomly choose samples for lens/otolith chronology
This code chunk randomly selects fish for chronology analysis. Lengths are grouped within 25-mm increments (i.e., LG-1 = 25-49mm, LG-2 = 50-74mm, LG-3 = 75-99, etc.). For each length group, 2 fish from each species are randomly sampled. If there was only 1 fish in a given species-length group, then only 1 fish was sampled.
```{r chronologysamples, include = FALSE}

set.seed(101)
samples1 <- lw22 %>% 
  group_by(species, lg_group) %>% 
  summarise("ID" = sample(ID, 2, replace = FALSE))
samples2 <- lw22 %>% 
  group_by(species, lg_group) %>%
  filter(length(lg_group) == 1) %>% 
  dplyr::select(species, lg_group, ID)
samples <- rbind(samples1, samples2)
samples <- samples %>% unite(spp_ID, c(species,ID), sep = "_", remove = FALSE)
lens.samples <- inner_join(dplyr::select(lw22, c(spp_ID, species, length_mm, lg_group)), dplyr::select(samples, c(species, spp_ID, lg_group)), by = c("spp_ID", "species", "lg_group"))

n.samples <- lens.samples %>% 
  group_by(species) %>% 
  summarise("n.samples" = length(unique(spp_ID)))
n.samples
# 23 ARCS, 35 BDWF, 16 HBWF, 21 LSCS

list.samples <- data.frame("Samples" = lens.samples$spp_ID, "Otolith" = NA, "Lens" = NA)
list.samples[is.na(list.samples)] <- ""
# kable(list.samples, caption = "Sample IDs for Otolith and Lens Chronology Analysis") %>% 
  # kable_styling(bootstrap_options = c("hover", "responsive"), full_width = F)

```

# Randomly choose samples for compound-specific SIA
```{r CSIAsamples, include = FALSE}
# Choose the largest individual from each species which were chosen for lens chronology analysis (plus ORNRC-tagged BDWF):
# BDWF_62, HBWF_111, ARCS_85, LSCS_140

csia.samples <- data.frame("spp_ID" = c("BDWF_62", "HBWF_111", "ARCS_85", "LSCS_140"))
```


# Randomly choose samples for bulk SIA
```{r bulkSIAsamples, include = FALSE}
lens.samples %>% 
  filter(length_mm > 200) %>% 
  group_by(species) %>% 
  summarise("N" = length(spp_ID))
# ARCS = 15, BDWF = 23, HBWF = 15, LSCS = 12
# Samples for bulk: ARCS = 14, BDWF = 22, HBWF = 14, LSCS = 11
# Samples needed for bulk: ARCS = 6, BDWF = 0, HBWF = 6, LSCS = 9

set.seed(213)
bulkSIA.ARCS <- lw22 %>% 
  filter(!spp_ID %in% lens.samples$spp_ID) %>% 
  filter(species == "ARCS") %>% 
  filter(length_mm > 200) %>% 
  summarise("spp_ID" = sample(spp_ID, 6, replace = FALSE))

set.seed(471)
bulkSIA.HBWF <- lw22 %>% 
  filter(!spp_ID %in% lens.samples$spp_ID) %>% 
  filter(species == "HBWF") %>% 
  filter(length_mm > 200) %>% 
  summarise("spp_ID" = sample(spp_ID, 6, replace = FALSE))

set.seed(628)
bulkSIA.LSCS <- lw22 %>% 
  filter(!spp_ID %in% lens.samples$spp_ID) %>% 
  filter(species == "LSCS") %>% 
  filter(length_mm > 200) %>% 
  summarise("spp_ID" = sample(spp_ID, 9, replace = FALSE))

bulkSIA.samples <- rbind(
  lw22 %>% filter(species %in% c("ARCS", "HBWF", "LSCS")) %>% filter(length_mm > 200) %>% filter(!spp_ID %in% csia.samples$spp_ID) %>% filter(spp_ID %in% lens.samples$spp_ID),
  lw22 %>% filter(spp_ID %in% bulkSIA.ARCS$spp_ID),
  lw22 %>% filter(spp_ID %in% bulkSIA.HBWF$spp_ID),
  lw22 %>% filter(spp_ID %in% bulkSIA.LSCS$spp_ID),
  lw22 %>% filter(spp_ID %in% sample(filter(lens.samples, spp_ID != "BDWF_62" & species == "BDWF" & length_mm > 200)$spp_ID, 20, replace = FALSE, set.seed(385)))
)

```

# Selecting more samples
More samples were selected to get the final sample numbers to 25 for each species.
```{r moresamps, include = FALSE}
set.seed(4976)
moresamps1 <- lw22 %>% 
  filter(!spp_ID %in% bulkSIA.samples$spp_ID) %>% 
  filter(length_mm > 200) %>% 
  filter(species == "ARCS") %>% 
  summarise("spp_ID" = sample(spp_ID, 5, replace = FALSE))

set.seed(5644)
moresamps2 <- lw22 %>% 
  filter(!spp_ID %in% bulkSIA.samples$spp_ID) %>% 
  filter(length_mm > 200) %>% 
  filter(species == "BDWF") %>% 
  summarise("spp_ID" = sample(spp_ID, 4, replace = FALSE))

set.seed(7651)
moresamps3 <- lw22 %>% 
  filter(!spp_ID %in% bulkSIA.samples$spp_ID) %>% 
  filter(length_mm > 200) %>% 
  filter(species == "HBWF") %>% 
  summarise("spp_ID" = sample(spp_ID, 5, replace = FALSE))

set.seed(9843)
moresamps4 <- lw22 %>% 
  filter(!spp_ID %in% bulkSIA.samples$spp_ID) %>% 
  filter(length_mm > 200) %>% 
  filter(species == "LSCS") %>% 
  summarise("spp_ID" = sample(spp_ID, 5, replace = FALSE))

moresampsID <- c(moresamps1$spp_ID, moresamps2$spp_ID, moresamps3$spp_ID, moresamps4$spp_ID)

moresamps.df <- lw22 %>% 
  filter(spp_ID %in% moresampsID)
moresamps.df
```

# Final list of samples for Bulk and CSIA
```{r totalsamps, include = FALSE}
totalsamps <- rbind(bulkSIA.samples, moresamps.df, filter(lw22, spp_ID == "BDWF_62"))
totalsamps <- totalsamps %>% 
  filter(!spp_ID %in% c("ARCS_120","ARCS_93","ARCS_131","HBWF_91","LSCS_54"))

n.samps.spp <- totalsamps %>% 
  group_by(species) %>% 
  summarize("N" = n())
n.samps.spp

final.csia.samps.df <- totalsamps %>% 
  arrange(desc(length_mm)) %>% 
  group_by(species) %>% 
  slice(1:5)
final.csia.samps.df

final.bulk.samps.df <- totalsamps %>% 
  filter(!spp_ID %in% final.csia.samps.df$spp_ID)
final.bulk.samps.df
```

```{r sampletables, include = TRUE}

kable(final.csia.samps.df, caption = "Final CSIA Samples") %>% 
  kable_styling(bootstrap_options = c("hover", "responsive"), full_width = F)

kable(final.bulk.samps.df, caption = "Final Bulk Samples") %>% 
  kable_styling(bootstrap_options = c("hover", "responsive"), full_width = F)
```


