---
title: "Stomach Content Analysis"
author: "Jonah Bacon"
date: '2022-11-22'
output: 
  html_document:
      toc: true
      toc_float:
        collapsed: false
        smooth_scroll: true
        
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)

## Load libraries
library(data.table)
library(here)
library(tidyverse)
library(knitr)
library(ggsci)
library(vegan)
library(kableExtra)

## Load data
load(here::here("C:/Users/Bacon/Downloads/beaufort-fish-trophic-analysis/data/stomach_content_workspace.RData"))

```

# Stomachs with prey

```{r stomachpreypresence, echo = FALSE}
kable(stomachs.w.prey) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)

stomachs.w.prey %>% 
  gather(key = "Presence", value = "Value", 3:4, na.rm = T) %>% 
ggplot(aes(x = year_collected, y = Value, fill = Presence, label = Value)) +
  geom_col() +
  geom_text(size = 5, position = position_stack(vjust = 0.5)) +
  facet_wrap(vars(Species)) +
  scale_fill_discrete(name = "Prey presence", labels = c("Absent", "Present"), guide = guide_legend(reverse = TRUE)) +
  ylab("Number of stomachs") +
  xlab("Year collected")
```

# Average fullness

```{r averagefullness, echo = FALSE}
kable(average.fullness) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)
```

# Average stomach weight

```{r averageweight, echo = FALSE}
kable(average.weight) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)

ggplot(prey.presence) +
  geom_boxplot(aes(x = year_collected, y = log(Total_contents_weight_g)), outlier.shape = NA) +
  geom_jitter(aes(x = year_collected, y = log(Total_contents_weight_g)), color = "black", alpha = 0.9) +
  facet_wrap(vars(Species)) +
  ylab("Log(Total stomach contents weight)") +
  xlab("Year collected")

ggplot(prey.presence) +
  geom_boxplot(aes(x = year_collected, y = Total_contents_weight_g), outlier.shape = NA) +
  geom_jitter(aes(x = year_collected, y = Total_contents_weight_g), color = "black", alpha = 0.9) +
  facet_wrap(vars(Species)) +
  scale_y_continuous(limits = c(0,5)) +
  ylab("Total stomach contents weight") +
  xlab("Year collected")

ggplot(prey.presence) +
  geom_histogram(aes(x = Total_contents_weight_g), col = "white") +
  facet_grid(rows = vars(Species), cols = vars(year_collected)) +
  scale_x_continuous(limits = c(0,5)) +
  scale_y_continuous(limits = c(0,14), breaks = seq(0,14,2)) +
  ylab("Stomachs count") +
  xlab("Total stomach contents weight")
```

# Frequency of occurrence

```{r freqoccurrence, echo = FALSE}
kable(frequency.of.occurrence) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)
```

# Rate of occurrence

```{r rateoccurrence, echo = FALSE}
kable(rate.of.occurrence) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)
```

# Constitution of diet by year

## Count
```{r dietconstitutioncount, echo = FALSE}
ggplot(df18, aes(x = year_collected, y = Count.percent)) +
  geom_col(aes(fill = Prey_group)) +
  facet_wrap(vars(Species), nrow = 2) +
  scale_fill_discrete(name = "Prey group") +
  xlab("Year collected") +
  ylab("Percent of diet by count")
```

## Weight
```{r dietconstitutionweight, echo = FALSE}
ggplot(df15, aes(x = year_collected, y = Wt.percent)) +
  geom_col(aes(fill = Prey_group)) +
  facet_wrap(vars(Species), nrow = 2) +
  scale_fill_discrete(name = "Prey group") +
  xlab("Year collected") +
  ylab("Percent of diet by weight")
```

## Relative percent
```{r dietconstitutionpercent, echo = FALSE}
ggplot(df24, aes(x = year_collected, y = Percent.percent)) +
  geom_col(aes(fill = Prey_group)) +
  facet_wrap(vars(Species), nrow = 2) +
  scale_fill_discrete(name = "Prey group") +
  xlab("Year collected") +
  ylab("Percent of diet by relative percent")

```

## Occurrence
```{r dietconstitutionoccurrence, echo = FALSE}
ggplot(df21, aes(x = year_collected, y = Occurrence.percent)) +
  geom_col(aes(fill = Prey_group)) +
  facet_wrap(vars(Species), nrow = 2) +
  scale_fill_discrete(name = "Prey group") +
  xlab("Year collected") +
  ylab("Percent of diet by frequency of occurrence")
```


# Constitution of diet by size

## Count
```{r dietconstitutioncount2, echo = FALSE}
ggplot(df6, aes(x = lg_group, y = Count.percent)) +
  geom_col(aes(fill = Prey_group)) +
  facet_grid(rows = vars(Species), cols = vars(year_collected)) +
  scale_fill_discrete(name = "Prey group") +
  scale_x_continuous(limits = c(3,11), breaks = seq(1:11)) +
  xlab("Length group (FL * 50)") +
  ylab("Percent of diet by count")
```

## Weight
```{r dietconstitutionweight2, echo = FALSE}
ggplot(df3, aes(x = lg_group, y = Wt.percent)) +
  geom_col(aes(fill = Prey_group)) +
  facet_grid(rows = vars(Species), cols = vars(year_collected)) +
  scale_fill_discrete(name = "Prey group") +
  scale_x_continuous(limits = c(1,11), breaks = seq(1:11)) +
  xlab("Length group (FL * 50)") +
  ylab("Percent of diet by weight")
```

## Relative percent
```{r dietconstitutionpercent2, echo = FALSE}
ggplot(df12, aes(x = lg_group, y = Percent.percent)) +
  geom_col(aes(fill = Prey_group)) +
  facet_grid(rows = vars(Species), cols = vars(year_collected)) +
  scale_fill_discrete(name = "Prey group") +
  scale_x_continuous(limits = c(1,11), breaks = seq(1:11)) +
  xlab("Length group (FL * 50)") +
  ylab("Percent of diet by relative percent")
```

## Occurrence
```{r dietconstitutionoccurrence2, echo = FALSE}
ggplot(df9, aes(x = lg_group, y = Occurrence.percent)) +
  geom_col(aes(fill = Prey_group)) +
  facet_grid(rows = vars(Species), cols = vars(year_collected)) +
  scale_fill_discrete(name = "Prey group") +
  scale_x_continuous(limits = c(1,11), breaks = seq(1:11)) +
  xlab("Length group (FL * 50)") +
  ylab("Percent of diet by frequency of occurrence")
```


# Diversity & Evenness
## Diversity indices

```{r diversityindices, include = FALSE}
kable(diversity.index.counts) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)

kable(diversity.index.percents) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)

kable(diversity.index.weights) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)
```

## Evenness index

```{r evennessindex, include = FALSE}
kable(H.max.counts) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)

kable(evenness.index.counts) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)

kable(H.max.weights) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)

kable(evenness.index.weights) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)

kable(H.max.percents) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)

kable(evenness.index.percents) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)
```

## Summarized Diversity/Evenness indices

```{r summarizeddiveven, echo = FALSE}
kable(summarized.diversity.evenness) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)
```


# Diet Overlap
## Prey count

```{r schoenercount1, include = FALSE}
kable(total.prey.count) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)
```

```{r schoenercount2, echo = FALSE}
kable(schoener.index.count) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)
```

## Prey weight

```{r schoenerweight1, include = FALSE}
kable(total.prey.weight) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)
```

```{r schoenerweight2, echo = FALSE}
kable(schoener.index.weight) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)
```

## Prey relative percent

```{r schoenerpercent1, include = FALSE}
kable(total.prey.percent) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)
```

```{r schoenerpercent2, echo = FALSE}
kable(schoener.index.percent) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)
```

## Summarized Schoener index

```{r summarizedschoener, echo = FALSE}
kable(combined.schoener.index) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F)
```
